---
title: "Project"
author: "Group5"
date: '2022-03-22'
output: pdf_document
---

GitHub Link: https://github.com/zazauwu/432_Group5

# Load Package
```{r}
library(ape)
library(tidyr)
library(MASS)
library(MuMIn)
library(ggplot2)
library(reshape2)
library(grid)
library(gridExtra)
library(ggfortify)
library(dplyr)

```

# Import Data
```{r}
###NK & HB lifetime reproductive success data
ltrsHB <- read.csv("./Data/HargreavesEcolLett data ltrsHB.csv")
ltrsNK <- read.csv("./Data/HargreavesEcolLett data ltrsNK.csv")

###NK & HB phenology data
phenNK <- read.csv("./Data/HargreavesEcolLett data phenologyNK.csv")
phenHB <- read.csv("./Data/HargreavesEcolLett data phenologyHB.csv")

###Climate by site data
clim <- read.csv("./Data/HargreavesEcolLett data hobo&ibutton summary by site.csv")
```

## Data Check and Modification
```{r}
sapply(clim, class)
```
```{r}
#variable explanation:
  #d.snowpack.10.11 = days with snowpack winter 2010-2011
  #wTmin.10.11 = minimum temperature winter (C) at plant height 2010-2011
  #Dec.FebTav.10.11 = mean temperature at plant height from Dec1st 2010 - Feb 28th 2011
  #d.grows.11 = # days in growing season 2011 (start / end of gorwing season determined by sustained snow or frost)
  #GDD.grows.11 = # growing degree days (Tbase=10C, Tmax capped at 30C) in growing season of 2011
  #GDD.yr.11 = # growing degree days in 2011 (ie not truncated to estimated growing season)
  #avmaxminTJuly.11 = mean July 2011 temperature (C) calculated from daily min and max T: 
      #mean(July1stmin + July1stmax + July2ndmin + July2ndmax + ... July31stmi + July31stmax)
  #avTJuly.11 = mean July 2011 temperature (C) calculated from raw data (ie >2 measurements per day)

#hence, convert the above columns into numeric variables
cols.num <- c("d.snowpack.11.12", "wTmin.11.12", "Dec.FebTav.11.12", "d.grows.12", "GDD.grows.12", "GDD.yr.12", "avmaxminTJuly.12", "avTJuly.12", "GDD.grows.13", "GDD.yr.13", "avmaxminTJuly.13", "avTJuly.13", "d.snowpack.13.14", "wTmin.13.14", "Dec.FebTav.13.14")
clim[cols.num] <- sapply(clim[cols.num],as.numeric)
sapply(clim, class)
head(clim)
```
```{r}
#use 'melt' to convert the 'clim' data.frame into long format in which 'transect' and 'site' are identifier variables, while the other are considered as the measured variables.
clim.sea <- melt(clim, id.vars = c('transect','site'), measure.vars = 3:29)
head(clim.sea)
```
```{r}
#use regular expression 'grepl' and 'gsub' to add 'year' column and remove date label in the 'variable' column
clim.sea$year <- ifelse(grepl('10.11', long$variable), 
                    'winter 10-11',
                    ifelse(grepl('11.12', long$variable), 
                           'winter 11-12',
                           ifelse(grepl('12.13', long$variable), 
                                  'winter 12-13',
                                  ifelse(grepl('13.14', long$variable), 
                                         'winter 13-14',
                                         ifelse(grepl('grows.11|July.11|yr.11', long$variable), 
                                                'year 2011',
                                                ifelse(grepl('grows.12|July.12|yr.12', long$variable), 
                                                       'year 2012',
                                                       ifelse(grepl('grows.13|July.13|yr.13', long$variable), 
                                                              'year 2013',
                                                              NA)))))))

clim.sea <- clim.sea %>%
  mutate(variable = gsub(".(\\d).*", "\\2", variable))

```

# Combine Datasets and Subset 
```{r}

# Merging HB and NK phenology datasets

phenmerged <- rbind(phenNK, phenHB)

# Subset phendat to variable we want

phendat <- phenmerged %>% dplyr::select(peak = transect, year, source = Source, site, treatment = treat, 
                                        final_fate = final.fate, DEm_plotavg = DEm.estplotav)



```




## PCA
Take phenology data and look at low v mid v high (Q2 and Q3) 
Are low similar to each other, are mid similar, are high similar 
Origin vs transplant location 
Phenotypic plasticity 
```{r}
# delete rows with NA value??
PhenoNA <- phenmerged %>%
    na.omit
```

```{r}
names(phenmerged)
```
```{r}
str(phenmerged)
```
```{r}
phenmerged$year <- as.factor(phenmerged$year)
row.names(phenmerged) <- phenmerged$X
```


Do plants transplanted from low-, mid-, and high- ranges vary in their ability to phenologically adapt to ranges outside their native range? (Fig. 2)
```{r}

```

